---
title: "EDA on pooled quad data"
output: github_document
editor_options: 
  chunk_output_type: console
---

# Setup

Load packages, etc.

```{r setup, include=T}
source("workflow/00_setup.R")
knitr::opts_knit$set(self.contained=T)
knitr::opts_chunk$set(echo = TRUE, cache = T, fig.width = 10, fig.height = 10)
library(sf)
```


```{r libraries, warning=FALSE,message=FALSE, cache=F}
library(foreach)
library(doParallel)
library(sf)
library(tidyverse)

library(viridis)
library(GGally)

```

## Data Import

This reads in the data object created with Data_Processing.Rmd.

```{r loaddata}
dataversion="20190128"
load(file.path("data",dataversion,"multi_quad_sample.Rdata"))
```
### Data Description

An incomplete description of columns in this dataset.

* `dist`: distance to the smoother surface but without a sign - so hills and holes are both positive
* `hole`: `dist`x`sign` so holes are negative and hills positive.
* `gc`: unsigned gaussian curvature with specified radius
* `gcs`: signed gaussian curvature (the sign was created from the dist metric)

See Data_Processing.Rmd for the code.




Gather data to 'long' (e.g. tidy) form for easier plotting with ggplot.
```{r meltdata}
d=poolquad

# Transform to 'long' format for easier ggplotting, etc.
dl=poolquad%>%
  ungroup()%>%
  gather(key=var,
                 value = val,-quad,-X,-Y,-Z,-id,-taxa,-class,
                 -classn,-rec,-pres,-contains("density"))%>%
  separate(var,c("var","scale"),sep="_")
```

# Exploratory Data Analysis

Plot the distribution of environmental variables.

## Correlation matrix

Correlation between the variables. Surprisingly even the two "rough" variables are very different.
```{r, warning=F}
my_bin <- function(data, mapping, ..., low = "#132B43", high = "red") {
  ggplot(data = data, mapping = mapping) +
    geom_hex(...) +
    scale_fill_gradient(low = low, high = high)
}

d%>%
  ungroup()%>%
  dplyr::select(contains("rough"),contains("hole"),contains("gcs"))%>%
  ggpairs(lower = list(continuous = my_bin))
```


## Box plot comparing background with presences

```{r, warning=F}
dl%>%
  ggplot(aes(x=pres,y=val))+
  geom_boxplot()+
  facet_grid(scale~var, scales="free")+
  xlab("Presence/Absence")+
  ylab("Environmental Variable")
```


```{r}
d%>%
  ggplot(aes(x=X,y=Y,col=Z))+
  facet_wrap(~quad,scales = "free")+
  geom_point(size=.1)+
  geom_point(data=filter(d,!is.na(rec)),aes(shape=taxa),col="red")+ #add the recruits
  scale_color_viridis_c()+
  coord_equal()

d%>%
  ggplot(aes(x=X,y=Y,col=Z_smooth_5))+
  facet_wrap(~quad,scales = "free")+
  geom_point(size=.1)+
  geom_point(data=filter(d,!is.na(rec)),aes(shape=taxa),col="red")+ #add the recruits
  scale_color_viridis_c()

```


```{r}
d%>%
  ggplot(aes(x=X,y=Y,col=slope_5))+
  facet_wrap(~quad,scales = "free")+
  geom_point(size=.1)+
  geom_point(data=filter(d,!is.na(rec)),aes(shape=taxa),col="red")+
  scale_color_viridis_c()

```

## Take a transect through the middle of the quad
```{r}
pick <- function(condition){ function(d) d %>% filter(!!enquo(condition)) }

ycut=0.02  #distance from ymean to include in the 'transect' 

# get mean y values to make a 'transect' through the middle
ymeans=d%>%
  group_by(quad)%>%
  summarize(ymean=mean(Y))%>%
  mutate(ymin=ymean-ycut,ymax=ymean+ycut)



d%>%
  ggplot(aes(x=X,y=Y,col=slope_5))+
  facet_wrap(~quad,scales = "free")+
  geom_point(size=.1)+
  geom_point(data=filter(d,!is.na(rec)),aes(shape=taxa),col="red")+
  scale_color_viridis_c()


d%>%
  left_join(ymeans,by="quad")%>%
  filter(abs(Y-ymean)<ycut)%>%
  ggplot(aes(x=X,y=Z,col=slope_5))+
  facet_wrap(~quad,scales = "free")+
  geom_point(size=.1)+
  geom_point(data= pick(~(!is.na(rec)&abs(Y-ymean)<ycut)),aes(shape=taxa),col="red")+
  scale_color_viridis_c()

```

## 2D histogram with location of recruits
```{r}
d%>%
  ggplot(aes(x=rough_20,y=slope_20))+
  facet_wrap(~quad,scales = "free")+
  geom_hex(bins=30)+
  geom_point(
    data=filter(d,!is.na(rec)),
    aes(shape=taxa),
    col="red")+
  scale_fill_viridis_c()
```


## Substrate Type

```{r plotSubstrate, warning=F}
d%>%
  ggplot(aes(x=as.factor(class),group=as.factor(pres),fill=as.factor(pres)))+
  stat_count(aes(y=..prop..), position='dodge')+
  scale_fill_manual(values=c("black","red"),name="Presence\nAbsence")+
  xlab("Point Classification")+
  ylab("Proportion of Presence/Absences")
```


## 3D Plot

Plot of the new 'hole' metric.

```{r,as.is=T, cache=F}
scatter3D(d$X, d$Y, d$Z,colvar = d$hole_10,bty = "g", cex=.01,theta = 100, phi = 0)
```
