---
title: "3D SDM Example"
output: github_document
editor_options: 
  chunk_output_type: console
---

# Setup

Load packages, etc.

```{r setup, include=T}
source("workflow/00_setup.R")
knitr::opts_knit$set(self.contained=T, root.dir="..")
knitr::opts_chunk$set(echo = TRUE, cache = T, fig.width = 10, fig.height = 10)
```


```{r libraries, warning=FALSE,message=FALSE, cache=F}
options(java.parameters = "-Xmx2048m")
library(rJava)
library(foreach)
library(doParallel)
library(rgdal)
library(sf)
library(dismo)
library(ENMeval)
library(ggplot2)
library(dplyr)
library(tidyr)

library(viridis)
library(GGally)

```

## Data Import

This reads in the data object created with Data_Processing.Rmd.

```{r loaddata}
load("data/model/multi_quad_sample.Rdata")

```
### Data Description

An incomplete description of columns in this dataset.

* `dist`: distance to the smoother surface but without a sign - so hills and holes are both positive
* `hole`: `dist`x`sign` so holes are negative and hills positive.
* `gc`: unsigned gaussian curvature with specified radius
* `gcs`: signed gaussian curvature (the sign was created from the dist metric)

See Data_Processing.Rmd for the code.




Gather data to 'long' (e.g. tidy) form for easier plotting with ggplot.
```{r meltdata}

# Transform to 'long' format for easier ggplotting, etc.
dl=poolquad%>%
  ungroup()%>%
  gather(key=var,
                 value = val,-X,-Y,-Z,-id,-taxa,-class,
                 -classn,-rec,-pres,-contains("density"))%>%
  separate(var,c("var","scale"),sep="_")
```

# Exploratory Data Analysis

Plot the distribution of environmental variables.

## Correlation matrix

Correlation between the variables. Surprisingly even the two "rough" variables are very different.
```{r, warning=F}
my_bin <- function(data, mapping, ..., low = "#132B43", high = "red") {
  ggplot(data = data, mapping = mapping) +
    geom_hex(...) +
    scale_fill_gradient(low = low, high = high)
}

d%>%
  ungroup()%>%
  dplyr::select(contains("rough"),contains("hole"),contains("gcs"))%>%
  ggpairs(lower = list(continuous = my_bin))
```


## Violin plot comparing background with presences

```{r, warning=F}
dl%>%
  ggplot(aes(x=presence,y=val))+
  geom_violin()+
  facet_grid(scale~var, scales="free_x")+
  xlab("Presence/Absence")+
  ylab("Environmental Variable")+
  scale_y_log10()+
  coord_flip()
```

## Substrate Type

```{r plotSubstrate, warning=F}
d%>%
  ggplot(aes(x=as.factor(class),group=as.factor(pres),fill=as.factor(pres)))+
  stat_count(aes(y=..prop..), position='dodge')+
  scale_fill_manual(values=c("black","red"),name="Presence\nAbsence")+
  xlab("Point Classification")+
  ylab("Proportion of Presence/Absences")
```


## 3D Plot

Plot of the new 'hole' metric.

```{r,as.is=T, cache=F}
scatter3D(d$X, d$Y, d$Z,colvar = d$hole_10,bty = "g", cex=.01,theta = 100, phi = 0)

# Or make an interactive plot:
#nColors <- 64
#cols <- as.character(cut(fulld$dists,breaks=nColors,labels= inferno(nColors)))
#scatterplot3js(fulld$x,fulld$y,fulld$z, color=cols,size = 0.05,stroke = NULL,pch=".")
```

# Distribution Model

## Prepare data

Create a subsetted, non-spatial table to use for model fitting and scale the variables.

```{r preparefitdata}
nbg=10000  # number of background points to select

subset_id=c(d$id[d$pres==1], # keep all presences
            d$id[sample(which(d$pres==0),nbg,replace = F)])  # sample only nbg background points

fitdata=d%>%
  mutate(pres=as.logical(pres),
         class=as.factor(class))%>%
  filter(id%in%subset_id)%>%
  select(pres,X,Y,Z,class,gcs_10,gcs_5,rough_10,rough_5,hole_10,hole_5,slope,aspect)%>% #select which variables to include
  mutate_if(is.numeric, function(x) as.vector(scale(x)))%>%  #scale all numeric variables so they have a mean of 0 and slope of 1
  na.omit()#%>% # drop rows with missing variables


```

## Fit Maxent Distribution Model

There are lots of options that will need to be considered.  This is one basic way:

```{r maxent}
m1 <- maxent(x=select(fitdata,-X,-Y,-Z,-pres),
             p=as.numeric(fitdata$pres),
    args=c(
		'randomtestpoints=30',
		'betamultiplier=1',
		'linear=true',
		'quadratic=true',
		'product=false',
		'threshold=false',
		'hinge=true',
		'threads=4',
		'responsecurves=true',
		'jackknife=true',
		'askoverwrite=false'))

```

## Validation

Need to add model comparison, validation, etc.

## Results

### Variable importance

```{r}
plot(m1)
```


### Response Curves

Relationship between each predictor (x-axis) and 'habitat suitability.'  These are estimates of the 'realized recuitment niche' based on the environmental data.  Don't put too much confidence in these results - we need to add more data and do much better model validation, etc.

```{r}
response(m1,expand=0)
```

### Predictions

Predict suitability across the full landscape (for each point).
```{r, warning=F}

# make random groups to reduce memory during predictions
d$group=rep(1:100,len=nrow(d))

# calculate the predictions for each point
p1=d%>%
  group_by(group)%>% #this breaks the full dataset down by group to make the predict() go faster
  do(pred=predict(m1, x=.))%>%  # predict habitat suitability for each point
  unnest()

```

## Predicted habitat suitability

3D plot of relative habitat suitability.  Areas in red are more suitable.

```{r}
#nColors <- 10
#cols <- as.character(cut(p1$pred,breaks=nColors,labels= rainbow(nColors)))

scatter3D(d$X, d$Y, d$Z,colvar = log(p1$pred+1),
          bty = "g", cex=.01,theta = 100, phi = 0)

```

Interactive 3D Figure of Habitat Suitability (same as above using different plotting function).  This will only work when you re-run it locally - not on github.

```{r,as.is=T, cache=F, eval=F}
scatterplot3js(fitdata$x,fitdata$y,fitdata$z, color=cols,size = 0.05,stroke = NULL,pch=".")
```

## Transect

Show values along a transect.

Extract a transect (I selected the y values using cloud compare).  This really shows the value of doing this in 3D - it can easily capture really complex terrain including caves, etc.

```{r}
transect=
  cbind.data.frame(d,pred=p1$pred)%>%
#  filter(between(y,-0.122,-0.121))%>%
  filter(between(Y,-0.027,-0.0265))%>%
  mutate(gcs_5=sign(gcs_5)*log(abs(gcs_5)),gcs_10=sign(gcs_10)*log(abs(gcs_10)))%>%
  gather("var","value",-X,-Y,-Z,-pres,-class,-pred,-taxa,-id)%>%
  arrange(pred,X,Z)
```

Environmental values:
```{r}
transect%>%
  filter(var%in%c("gcs_10","gcs_5"))%>%
  ggplot(aes(x=X,y=Z,col=value))+
  geom_point(size=.7)+
  scale_color_gradient2(low="blue",mid="grey",high="red",name="Variable\nValue")+
  facet_wrap(~var,ncol=1)+
  xlab("Distance along transect (m)")+
  ylab("Value")+
  coord_equal()

transect%>%
  filter(var%in%c("hole_10","hole_5"))%>%
  ggplot(aes(x=X,y=Z,col=value))+
  geom_point(size=.7)+
  scale_color_gradient2(low="blue",mid="grey",high="red",name="Variable\nValue")+
  facet_wrap(~var,ncol=1)+
  xlab("Distance along transect (m)")+
  ylab("Value")+
  coord_equal()

transect%>%
  filter(var%in%c("rough_10","rough_5"))%>%
  ggplot(aes(x=X,y=Z,col=value))+
  geom_point(size=.7)+
  scale_color_gradient2(low="blue",mid="grey",high="red",name="Variable\nValue")+
  facet_wrap(~var,ncol=1)+
  xlab("Distance along transect (m)")+
  ylab("Value")+
  coord_equal()
```

Predicted habitat suitability.  Note some (but not all) little holes full of red.
```{r}
ggplot(transect,aes(x=X,y=Z,col=pred))+
  geom_point(size=.3)+
  scale_color_gradientn(colors=c("blue","yellow","red"),
                        name="Habitat\nSuitability",trans="log1p")+
  xlab("Distance along transect (m)")+
  ylab("Height")+
  coord_equal()
```

An added bonus is that this transect kind of looks like a buffalo (head with horns on the left side).  It looks a little messy there because the transect is actually a little ribbon (not just a one-dimensional line) and so includes several values at each y position.  We can think of better ways to show that.

## Export data

Export the data (with predictions) in a format that can be opened by cloud compare.

```{r, eval=F}
write.csv(cbind.data.frame(d,pred=p1$pred),"data/modeloutput.txt")
```

## Render this document to markdown
```{r, eval=F}
library(ezknitr)
ezknit(file = "vignettes/SDM_3D.Rmd", out_dir = "vignettes")
```
